---
title: "FB ADS"
author: "Ondrej Pekacek/TI CZ"
date: "Updated `r format(Sys.time(),'%d. %m. %Y')`"
output: 
  flexdashboard::flex_dashboard:
    logo: "data/logo_ti.png"
    theme: cosmo
    orientation: columns
    vertical_layout: fill
    source_code: https://github.com/opop999/election_monitoring_fb_ads
    navbar:
      - {title: "Project", icon: "ion-information-circled", href: "https://www.transparentnivolby.cz/snemovna2021"}
      - {title: "Author", icon: "ion-social-linkedin", href: "https://www.linkedin.com/in/ondrej-pekacek"}
      - {title: "Data: FB", icon: "ion-cloud", href: "https://www.facebook.com/ads/library/"}
---

```{r setup, include=FALSE}
# Disable scientific notation of numbers
options(scipen = 999)

# Package names
packages <- c("flexdashboard", "dplyr", "ggplot2", "plotly", "forcats", "htmlwidgets", "tidyr")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

# Specify directory with datasets
directory_datasets <- "data/temp"

# Import datasets created in the previous script
summary_dataset <- readRDS(file = paste0(directory_datasets, "/summary_tables/merged_summary.rds"))
time_dataset <- readRDS(file = paste0(directory_datasets, "/summary_tables/time_summary.rds"))

# Specify output directory for individual plots
directory <- "data/temp/html_plots"

# Check whether output directory exists to save individual plots
if (!dir.exists(directory)) {
  dir.create(directory)
} else {
  print("Output directory already exists")
}

# Select TOP 10 biggest spenders for plotting
# TEMPORARY SUBFILTER:
# Filter out entities, which are not current 2021 election candidates
top_10_spenders <- summary_dataset %>%
  filter(!page_id %in% c(102389958091735, 728495140691300)) %>% 
  slice_max(avg_spend, n = 10) %>%
  select(page_name, page_id, avg_spend)

# Election date for vertical line in the time-plots
election_date <- as.Date("2021-10-08")

# Graph zoom date end & beginning
start_date <- as.Date("2021-01-01")
end_date <- as.Date("2021-11-05")

```

Ads
=====================================

Column
-----------------------------------------------------------------------

### Total Ads

```{r}
plot_total_ads <- ggplotly(summary_dataset %>%
  filter(total_ads > 10) %>%
  ggplot(aes(x = total_ads, y = reorder(page_name, total_ads), fill = percent_unique)) +
  geom_col() +
  scale_fill_gradient(
    low = "#fbe8e7", high = "#db1d0b",
    limits = c(0, 1),
    breaks = c(0, 0.25, 0.50, 0.75, 1),
    labels = c("0%", "25%", "50%", "75%", "100%")
  ) +
  scale_x_continuous(
    breaks = seq(0, 2500, 250),
    labels = seq(0, 2500, 250)
  ) +
  theme_minimal() +
  xlab("Total Ads") +
  ylab(element_blank()) +
  labs(fill = "Proportion unique ads") +
  ggtitle(paste("Total Ads (pages with more than 10 ads) since", format(start_date, "%d.%m.%Y"))))

plot_total_ads

saveWidget(plot_total_ads, file = paste0(directory, "/plot_total_ads.html"))

```

Spending
=====================================

Column 
-----------------------------------------------------------------------

### Total Spending 

```{r}
plot_total_spending <- ggplotly(summary_dataset %>%
  filter(total_ads > 10) %>%
  mutate(
    avg_spend_1000 = round(avg_spend / 1000, digits = 0),
    per_ad_avg_spend_1000 = round(per_ad_avg_spend / 1000, digits = 1)
  ) %>%
  ggplot(aes(x = avg_spend_1000, y = reorder(page_name, avg_spend), fill = per_ad_avg_spend_1000)) +
  geom_col() +
  scale_fill_gradient2(
    low = "#cddae5",
    mid = "#03457f",
    high = "#023766",
    midpoint = 12,
    limits = c(0, 25)
  ) +
  scale_x_continuous(
    breaks = seq(0, 3000, 250),
    labels = seq(0, 3000, 250)
  ) +
  theme_minimal() +
  xlab("Total Spend (CZK thousands)") +
  ylab(element_blank()) +
  labs(fill = "Per Ad (CZK thousands)") +
  ggtitle(paste("Spending on Ads since", format(start_date, "%d.%m.%Y"))))

plot_total_spending

saveWidget(plot_total_spending, file = paste0(directory, "/plot_total_spending.html"))

```

Runtime
=====================================

Column 
-----------------------------------------------------------------------

### Ad Runtime 

```{r}
plot_total_runtime <- ggplotly(summary_dataset %>%
  filter(total_ads > 10) %>%
  ggplot(aes(x = avg_ad_runtime, y = reorder(page_name, avg_ad_runtime))) +
  geom_col(fill = "#026663") +
  scale_x_continuous(
    breaks = seq(0, 30, 2),
    labels = seq(0, 30, 2)
  ) +
  theme_minimal() +
  xlab("Days") +
  ylab(element_blank()) +
  ggtitle(paste("Average Ad runtime since", format(start_date, "%d.%m.%Y"))))

plot_total_runtime

saveWidget(plot_total_runtime, file = paste0(directory, "/plot_total_runtime.html"))

```

Impressions
=====================================

Column 
-----------------------------------------------------------------------

### Total Impressions (square root scale)

```{r}
plot_total_impressions <- ggplotly(summary_dataset %>%
  filter(total_ads > 10) %>%
  mutate(
    avg_impressions_million = round(total_avg_impressions / 1000000, digits = 3),
    per_ad_avg_impressions_1000 = round(per_ad_avg_impression / 1000, digits = 1)
  ) %>%
  ggplot(aes(x = avg_impressions_million, y = reorder(page_name, avg_impressions_million), fill = per_ad_avg_impressions_1000)) +
  geom_col() +
  scale_fill_gradient2(
    low = "#e7fbe8",
    mid = "#089914",
    high = "#011603",
    midpoint = 200,
    limits = c(0, 400)
  ) +
  scale_x_sqrt(
    breaks = seq(0, 100, 10),
    labels = seq(0, 100, 10)
  ) +
  theme_minimal() +
  xlab("Total Impressions (millions)") +
  ylab(element_blank()) +
  labs(fill = "Per Ad impressions (thousands)") +
  ggtitle(paste("Total Ads impressions (square root scale) since", format(start_date, "%d.%m.%Y"))))

plot_total_impressions

saveWidget(plot_total_impressions, file = paste0(directory, "/plot_total_impressions.html"))

```

Reach
=====================================

Column 
-----------------------------------------------------------------------

### Total Reach (minimal, square root scale)

```{r}
plot_total_reach <- ggplotly(summary_dataset %>%
  filter(total_ads > 10) %>%
  mutate(
    avg_min_reach_million = round(total_min_reach / 1000000, digits = 2),
    per_ad_avg_min_reach_1000 = round(per_ad_min_reach / 1000, digits = 0)
  ) %>%
  ggplot(aes(x = avg_min_reach_million, y = reorder(page_name, avg_min_reach_million), fill = per_ad_avg_min_reach_1000)) +
  geom_col() +
  scale_fill_gradient2(
    low = "#f0e6eb",
    mid = "#660237",
    high = "#14000b",
    midpoint = 500,
    limits = c(0, 1000)
  ) +
  scale_x_sqrt(
    breaks = seq(0, 5000, 250),
    labels = seq(0, 5000, 250)
  ) +
  theme_minimal() +
  xlab("Total reach (millions)") +
  ylab(element_blank()) +
  labs(fill = "Per Ad reach (thousands)") +
  ggtitle(paste("Total minimal reach of Ads (square root scale) since", format(start_date, "%d.%m.%Y"))))

plot_total_reach

saveWidget(plot_total_reach, file = paste0(directory, "/plot_total_reach.html"))

```

Spending trend
=====================================

Column 
-----------------------------------------------------------------------

### Spending in Time (10 biggest spenders)

```{r}
spend_over_time <- ggplotly(
  time_dataset %>%
    filter(page_id %in% top_10_spenders$page_id) %>%
    mutate(
      ad_creation_time_week = as.Date(cut(ad_creation_time, breaks = "week", start.on.monday = TRUE)) + 4,
      cumulative_spend_million = round((cumulative_spend / 1000000), digits = 3)
    ) %>%
    group_by(page_name, ad_creation_time_week) %>%
    summarise(end_of_week_million_spend = max(cumulative_spend_million)) %>%
    ungroup() %>%
    ggplot(aes(x = ad_creation_time_week, y = end_of_week_million_spend, color = page_name)) +
    geom_line() +
    geom_point(size = 0.8) +
    geom_vline(aes(xintercept = as.numeric(election_date)), color = "#db1d0b") +
    geom_text(aes(x = election_date, y = 0.2, label = "elections"), color = "#03457f", size = 4) +
    theme_minimal() +
    scale_y_continuous(
      breaks = seq(0, 5, 0.25),
      labels = seq(0, 5, 0.25)
    ) +
    scale_x_date(date_breaks = "1 months", date_labels = "%d.%m.%y") +
    coord_cartesian(xlim = c(start_date, end_date), expand = TRUE) +
    theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1)) +
    xlab(element_blank()) +
    ylab("CZK million") +
    ggtitle(paste("Spending (in CZK million) on Ads since", format(start_date, "%d.%m.%Y")))
)


spend_over_time

saveWidget(spend_over_time, file = paste0(directory, "/plot_spend_over_time.html"))

```

Spending weekly
=====================================

Column 
-----------------------------------------------------------------------

### Spending Weekly (10 biggest spenders)

```{r}
spend_weekly <- ggplotly(
  time_dataset %>%
    filter(page_id %in% top_10_spenders$page_id) %>%
    mutate(
      end_of_week = as.Date(cut(ad_creation_time, breaks = "week", start.on.monday = TRUE)) + 4,
    ) %>%
    group_by(page_name, end_of_week) %>% 
    summarise(by_week_thousands_spend = round(sum(avg_spend) / 1000, digits = 0)) %>% 
    ungroup() %>% 
    ggplot(aes(x = end_of_week, y = by_week_thousands_spend, fill = page_name)) +
    geom_col(width = 5.5) +
    geom_vline(aes(xintercept = as.numeric(election_date)), color = "#db1d0b") +
    geom_text(aes(x = election_date, y = 0.2, label = "elections"), color = "#03457f", size = 4) +
    theme_minimal() +
    scale_y_continuous(
      breaks = seq(0, 5000, 200),
      labels = seq(0, 5000, 200)
    ) +
    scale_x_date(date_breaks = "1 months", date_labels = "%d.%m.%y") +
    coord_cartesian(xlim = c(start_date, end_date), expand = TRUE) +
    theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1)) +
    xlab(element_blank()) +
    ylab("CZK thousand") +
    ggtitle(paste("Weekly spending (in CZK thousand) on Ads since", format(start_date, "%d.%m.%Y")))
)

spend_weekly

saveWidget(spend_weekly, file = paste0(directory, "/plot_spend_weekly.html"))

```

Gender
=====================================

Column 
-----------------------------------------------------------------------

### Ads by Gender // **NOTE: DUE TO THE TECHNICAL DIFFICULTIES, DATA REGARDING DEMOGRAPHIC AND REGIONAL TARGETING OF POLITICAL ADS WERE LAST UPDATED ON 18.8.2021**

```{r include=FALSE}
# We reload the currently non-updated dataset with demographic and regional data
summary_dataset <- readRDS(file = "data/summary_tables/merged_summary.rds")

```

```{r}
plot_gender <- ggplotly(
  summary_dataset %>%
    filter(total_ads > 10) %>%
    ggplot(aes(x = avg_male, y = reorder(page_name, avg_male), fill = avg_male)) +
    geom_col() +
    geom_vline(aes(xintercept = 0.5, color = "#db0b50")) +
    scale_fill_gradient(
      low = "#db1d0b", high = "#03457f",
    ) +
    scale_x_continuous(
      limits = c(0, 1),
      breaks = seq(0, 1, 0.25),
      labels = c("0%", "25%", "50%", "75%", "100%")
    ) +
    theme_minimal() +
    theme(legend.position = "none") +
    xlab("Proportion Ads Targeting Male Audience") +
    ylab(element_blank()) +
    ggtitle(paste("Proportion of Ads targeting male audience since", format(start_date, "%d.%m.%Y")))
)

plot_gender

```

Age
=====================================

Column 
-----------------------------------------------------------------------

### Ads by Age (10 biggest spenders) // **NOTE: DUE TO THE TECHNICAL DIFFICULTIES, DATA REGARDING DEMOGRAPHIC AND REGIONAL TARGETING OF POLITICAL ADS WERE LAST UPDATED ON 18.8.2021**

```{r}
plot_age <- ggplotly(
  summary_dataset %>%
    filter(page_id %in% top_10_spenders$page_id) %>%
    select(
      page_name,
      avg_18_24,
      avg_25_34,
      avg_35_44,
      avg_45_54,
      avg_55_64,
      avg_65_plus
    ) %>%
    pivot_longer(cols = 2:7, names_to = "age_group", values_to = "proportion", values_drop_na = TRUE) %>%
    ggplot(aes(x = age_group, y = proportion, fill = page_name)) +
    geom_col(position = "dodge2") +
    scale_x_discrete(labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+")) +
    scale_fill_brewer(palette = "Paired") +
    scale_y_continuous(
      limits = c(0, 0.5),
      breaks = seq(0, 0.5, 0.1),
      labels = c("0%", "10%", "20%", "30%", "40%", "50%")
    ) +
    theme_minimal() +
    theme(legend.title = element_blank()) +
    xlab(element_blank()) +
    ylab(element_blank()) +
    ggtitle(paste("Proportion of Ads targeting specific age group since", format(start_date, "%d.%m.%Y")))
)

plot_age

```

Region
=====================================

Column 
-----------------------------------------------------------------------

### Ads by Region (10 biggest spenders) // **NOTE: DUE TO THE TECHNICAL DIFFICULTIES, DATA REGARDING DEMOGRAPHIC AND REGIONAL TARGETING OF POLITICAL ADS WERE LAST UPDATED ON 18.8.2021**

```{r}
plot_region <- ggplotly(
  summary_dataset %>%
    filter(page_id %in% top_10_spenders$page_id) %>%
    select(
      page_name,
      avg_pha,
      avg_stc,
      avg_jhc,
      avg_plk,
      avg_kvk,
      avg_ulk,
      avg_lbk,
      avg_hkk,
      avg_pak,
      avg_vys,
      avg_jhm,
      avg_olk,
      avg_msk,
      avg_zlk
    ) %>%
    pivot_longer(cols = 2:15, names_to = "region_group", values_to = "proportion", values_drop_na = TRUE) %>%
    ggplot(aes(x = region_group, y = proportion, fill = page_name)) +
    geom_col(position = "dodge2") +
    scale_x_discrete(labels = c("HKK", "JHC", "JHM", "KVK", "LBK", "MSK", "OLK", "PAK", "PHA", "PLK", "STC", "ULK", "VYS", "ZLK")) +
    scale_fill_brewer(palette = "Paired") +
    scale_y_continuous(
      limits = c(0, 0.5),
      breaks = seq(0, 0.5, 0.1),
      labels = c("0%", "10%", "20%", "30%", "40%", "50%")
    ) +
    theme_minimal() +
    theme(legend.title = element_blank()) +
    xlab(element_blank()) +
    ylab(element_blank()) +
    ggtitle(paste("Proportion of Ads targeting specific region since", format(start_date, "%d.%m.%Y")))
)

plot_region

```

```{r cleanup, include=FALSE}
# Because the saveWidget function does not correctly delete the dependency files
# which are used to create individual self-sustaining widgets, we have to delete
# them using R functions. All non-html files in output folder are deleted.

unlink(grep(
  x = list.files(
    path = directory,
    recursive = TRUE,
    full.names = TRUE
  ),
  pattern = "(.html)$",
  invert = TRUE,
  value = TRUE
))

```

